### 2015-12-18 ldfaxis2.tex
Refactored intro a bit

% Linked Data Fragments axis
\begin{figure}[t]
  \begin{tikzpicture}[
    x=.005\textwidth,
    y=.1cm,
    axis/.style={
      line width=.8pt,
      stealth-stealth,
      color=black!65,
    },
    axislabel/.style={
      inner sep=0,
      font=\slshape\fontsize{8}{9}\selectfont,
      color=black!70,
    },
    tick/.style={
      line width=1pt,
    },
    ticklabel/.style={
      anchor=south,
      inner sep=0,
      align=center,
      font=\bfseries\fontsize{8}{8}\selectfont,
      text depth=0pt,
    },
  ]
    \newcommand\tick[2]{%
      \draw[tick] (#1, 1.7) -- (#1, -.4pt);
      \node[ticklabel] at (#1, 2.4) {#2};
    }
    \newcommand\vaguetick[1]{%
      \draw[tick,densely dotted] (#1, 1.7) -- (#1, -.4pt);
    }

    \draw[axis] (-100, 0) -- (0, 0);
    \node[axislabel,align=left,anchor=north west] at (-100, -2.3)
    {
%      generic requests\\
      high client effort\\
      high server availability\\
      know how many times datadump was downloaded
    };
    \node[axislabel,align=right,anchor=north east] at (0, -2.3)
    {
%      specific requests\\
      high server effort\\
      low server availability\\
      detailled query logs
    };

    \tick{-93}{GTFS data\\dump}
    \tick{-10}{route planning\\service}
    \tick{-50}{Linked\\Connections}

    \vaguetick{-84}    \vaguetick{-70}
    \vaguetick{-10}
    \vaguetick{-28}
    \vaguetick{-20}

   %\node[align=center,font=\fontsize{8}{9}\selectfont\bfseries] at (0, -8.3)
   %   {various types of\\Linked Data Fragments};
  \end{tikzpicture}
  \caption{This axis extends \cref{fig:LDFAxis1} with Linked Connections: a different choice of trade-offs for route planning advice over the Web.
  }
  \label{fig:LDFAxis2}
\end{figure}